TCPTokens: Introducing Currency into Data Center Congestion Control

### 2017-10-10 Proposal/implementation.tex
Push near final draft for tonight

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation}
\label{sec:implementation}
We intend to emulate our system in Mininet to observe traffic patterns and infer a suitable token algorithm. Mininet has proven itself to be a viable tool to model new congestion control algorithms CITATION, and will help us prototype our concept efficiently. We will build a custom SDN controller that interacts with traditional OpenFlow software switches as well as end-hosts. End-hosts will run a custom real-world traffic generation script that adjust based on information packets send by the controller. 
If time permits, we may expand our implementation to MaxiNet, which can emulate large-scale network stress tests of thousands of nodes on multiple physical hosts.

We initially considered a second implementation alternative in C/C++ based on the FastPass  CITATION source code. This would provide us with a fully deployable system which we could fork our implementation from. A major advantage of this approach is the ability to test scenarios and traffic algorithms using real software code. 
However, several concerns made us favour a Mininet emulation instead.
First, FastPass relies on DPDK integration, which requires actual hardware interfaces. The central arbiter in the FastPass design would need to run on a dedicated machine, which increases prototyping and development complexity substantially.
Secondly, the FastPass code is highly specialized and optimized research code with only little available documentation. Modifying and evolving the source code will require thorough understanding of kernel and networking development, a significant time-sink. For a class project, these may be major initial hurdles, taking away from the research aspect of the design concept.
Consequently, we have decided to pursue an approach which allows us to quickly develop an understanding of the problem without being obstructed by engineering work.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation}
\label{sec:implementation}
We intend to emulate our system in Mininet~\cite{mininet} to observe traffic patterns and infer a suitable token algorithm. Mininet has proven itself to be a viable tool to model new congestion control algorithms~\cite{mininet_learning}, and will help us prototype our concept efficiently. We will build a custom SDN controller that interacts with traditional OpenFlow software switches as well as end-hosts. End-hosts will run a custom real-world traffic generation script which adjusts based on information packets sent by the controller. 
If time permits, we may expand our implementation to MaxiNet, which can emulate large-scale network stress tests of thousands of nodes on multiple physical hosts.

We initially considered a second implementation alternative in C/C++ based on the FastPass~\cite{fastpass} source code. This would provide us with a fully deployable system which we could fork our implementation from. A major advantage of this approach is the ability to test scenarios and traffic algorithms using real software code. 
However, several concerns made us favour a Mininet emulation instead.
Firstly, FastPass relies on DPDK integration, which requires actual hardware interfaces. The central arbiter in the FastPass design would need to run on a dedicated machine, which increases prototyping and development complexity substantially.
Secondly, the FastPass code is highly specialized and optimized research code with only little available documentation. Modifying and evolving the source code will require thorough understanding of kernel and networking development, a significant time-sink. For a class project, these may be major initial hurdles, taking away from the research aspect of the design concept.
Consequently, we have decided to pursue an approach which allows us to quickly develop an understanding of the problem without being obstructed by engineering work.

### 2017-10-10 Proposal/intro.tex
Push near final draft for tonight

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}

Initially, research in network congestion  was dominated by the  assuming a decentralised, autonomous network. End-hosts only have control over the amount of traffic they send, and are unaware of the intentions or traffic rates of their peers. Similarly, switches and routers are unaware of global traffic patterns and only forward based on their local notion of optimality.

In line with these assumptions, TCP was designed to optimize traffic globally on a simple fairness principle; nodes react to packet loss and assume that others will behave accordingly. An equilibrium is reached when all end-hosts achieve a traffic rate that, in sum, conforms to the maximum available bandwidth of the most congested link.
TCP works well  in scenarios where many different distrustful participants compete for limited bandwidth. Still, TCP is a \textit{reactive protocol}; the fact that packet loss and latency increases occur in the network already indicates a problem. Packet loss is primarily  due to overflowing  queues in forwarding elements, implying that traffic has not been optimally distributed. 
Ideally, a network should always be “zero-queue”, i.e., latency will merely be induced by propagation, and not queuing delay.

Queueing has generally not been a dominating issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few “hot” target hosts.~\cite{hedera, microte} Traffic optimization is a substantial challenge; network operators have no control over the individual network elements nor its participants. Under these conditions, TCP and its extensions can be considered a best-effort solution.

Developments in the past decade have changed the general networking environment. Datacenters have emerged as an exciting new research frontier, posing novel  design challenges and opportunities.
Driven by minimization of costs and maximization of compute power;  data centers must  run at maximum utilization to achieve an optimal compute/cost ratio. Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load network, requiring more sophisticated approaches to solve congestion control. 
On the other hand, operators now have the ability to freely control and adapt their network architecture leading highly customized systems and fine-grained optimization. As a result  Software-Defined Networking (SDN) emerged as a new networking paradigm. Moving away from the principle of distributed communication and routing, SDN introduces the notion of “centralized management”. A single controller with global knowledge is able to automatically modify and adapt the forwarding tables of all switches in the network, as well as notify end hosts of changes in the network.
These two new trends in systems facilitated impactful new innovation opportunities in the space of TCP congestion research. Traffic can now be managed in a  \textit{centralised} fashion based on \textit{global knowledge} of the entire topology and traffic patterns.
A new line of centralised schedulers has emerged that  can achieve close to optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
However, these schedulers are still  \textit{reactive}  in nature. The central controller responds to changes in the network or requests by applications, which may cost valuable roundtrip latency. Often, short-term flows or bursts are unaccounted for, which causes undesirable packet loss and backpropagating congestion.

A much more desirable solution is a global, centralised arbiter which is able to predict and fairly distribute flows in the network before bursts or congestion occurs. By treating the network’s compute and forwarding power as a single finite resource, a controller could act like the OS scheduler distributing CPU time slices to processes. This design approach also follows SDN’s aspiration of introducing operating systems abstractions to the networking domain space.

@@

In the scope of this course, we attempt to answer question 1 and design a simple token-based scheduler in Mininet. If we succeed, we will benchmark our results and evaluate the level of utilization compared to contemporary scheduling systems.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
Initially, research in network congestion was dominated by the assumption of a decentralised, autonomous network - end-hosts only have control over the amount of traffic they send, and are unaware of the intentions or traffic rates of their peers. Similarly, switches and routers are unaware of global traffic patterns and only forward based on their local notion of optimality.

In line with these assumptions, TCP was designed to optimize traffic globally on a simple fairness principle: nodes react to packet loss and assume that others will behave accordingly. An equilibrium is reached when all end-hosts achieve a traffic rate that, in sum, conforms to the maximum available bandwidth of the most congested link.
TCP works well  in scenarios where many different distrustful participants compete for limited bandwidth. Still, TCP is a \textit{reactive protocol}; the fact that packet loss and latency increases occur in the network already indicates a problem. Packet loss is primarily due to overflowing queues in forwarding elements, implying that traffic has not been optimally distributed. Ideally, a network should always be "zero-queue", i.e., latency will merely be induced by propagation, and not queuing delay.

Queueing has generally not been a dominating issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few "hot" target hosts.~\cite{hedera, microte} Traffic optimization is a substantial challenge; network operators have no control over the individual network elements nor its participants. Under these conditions, TCP and its extensions can be considered a best-effort solution.

Developments in the past decade have changed the general networking environment. Datacenters have emerged as an exciting new research frontier, posing novel  design challenges and opportunities.
Driven by minimization of costs and maximization of compute power, data centers must  run at maximum utilization to achieve an optimal compute/cost ratio. Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load network, requiring more sophisticated approaches to solve congestion control. 
On the other hand, operators now have the ability to freely control and adapt their network architecture, leading to highly customized systems and fine-grained optimization. As a result , Software-Defined Networking (SDN) has emerged as a new networking paradigm. Moving away from the principle of distributed communication and routing, SDN introduces the notion of "centralized management". A single controller with global knowledge is able to automatically modify and adapt the forwarding tables of all switches in the network, while notifying end hosts of changes in the network.
These two new trends in system design facilitated impactful new innovation opportunities in the space of TCP congestion research. Traffic can now be managed in a  \textit{centralised} fashion based on \textit{global knowledge} of the entire topology and traffic patterns.
A new line of centralised schedulers has emerged that  can achieve close to optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
However, these schedulers are still  \textit{reactive}  in nature. The central controller responds to changes in the network or requests by applications, which may cost valuable roundtrip latency. Often, short-term flows or bursts are unaccounted for, which causes undesirable packet loss and backpropagating congestion.

A much more desirable solution is a global, centralised arbiter which is able to predict and fairly distribute flows in the network before bursts or congestion occurs. By treating the network’s compute and forwarding power as a single finite resource, a controller acts like the OS scheduler distributing CPU time slices to processes. This design approach follows SDN’s aspiration of introducing operating systems abstractions to the networking domain space.

@@

In the scope of this course, we attempt to answer question 1 and design a simple token-based scheduler in Mininet. If successful, we will benchmark our results and evaluate the level of utilization compared to contemporary scheduling systems.

### 2017-10-10 Proposal/paper.tex
Push near final draft for tonight

\begin{abstract}

\end{abstract}

@@

\input{relatedworks}
\input{design}
%\input{motivation}
\input{implementation}
\input{eval}

>>>>

\begin{abstract}
Datacenter networks have become a hotbed for research in the recent past. Many works have been published on improving the bisection bandwidth, reducing cost, and improving congestion control as well as latency. As a core contribution of this research, centralized scheduling has entered the stage as dominant strategy to manage flows. Benefitting from a global view and full control over the network, centralized schedulers are able to enforce fine-grained traffic control, frequently achieving near-optimal bandwidth optimization. However, modern congestion control algorithms and schedulers are still fundamentally reactive. Many algorithms are designed to only respond to packet loss or significant increase in latency, not to actively prevent it.
In this project, we plan to explore the opportunities of employing a preemptive and tightly controlling central network scheduler. Using tokens as a global tool to enforce traffic limiting and policies, the scheduler is able to proactively control the flow of traffic as well as respond to application requests. "Knowledge" and predictive analysis in networks is a growing trend in research which we intend to leverage in this system.
We investigate the potential in such a new form of interactive congestion control and analyze it against to state-of-the-art solutions. Our tool of choice to model our system is Mininet, a rapid prototyping emulator for datacenter networks. We will compare our "TCPToken" system against established TCP-congestion systems such as Hedera and DCTCP. Based on the findings and measurement results, we will reassess the feasibility and prospect of a token-based, predictive congestion control algorithm.
\end{abstract}

@@

\input{relatedworks}
\input{design}
%\input{motivation}
\input{implementation}
\input{eval}
\input{timeline}
